---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# USpopcenters

<!-- badges: start -->
<!-- badges: end -->

USpopcenters contains data on the centers of population (also known as "centroids") of census areas in the United States.

Centers of population are according to the 2000 and 2010 decennial censuses. They are given via **longitude and latitude** coordinates and are available at the level of **state**, **county**, **tract**, and **block group**. The mean and median centers of population of the entire US according to the 2010 census are also included in this package.

All these data are available in this package in the form of [`tibble`s](https://github.com/tidyverse/tibble). Each year/geography combination has its own data set, and each data set also includes a column with each area's total population. The `tibble`s are indexed by GEOID/FIPS code, and some of the state-level and county-level tables also contain state and county names.

All applicable tables contain data from the fifty states and the District of Columbia. Non-state island territories' data has inconsistent availability from data set to data set. See each `tibble`'s documentation page to see what each one actually contains. 

This package was created primarily because the US Census Bureau has not made these data available via API. They were downloaded manually from the Census Bureau website at <https://www.census.gov/geographies/reference-files/time-series/geo/centers-population.html>.

## Installation

You can install the released version of USpopcenters from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("USpopcenters")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("NikKrieger/USpopcenters")
```
## Example

County-level centers of population from the 2010 US census:

```{r load_tibble, include=F}
library(tibble)
```

```{r example}
library(USpopcenters)
county2010
```

## US census data terms of service

<https://www.census.gov/data/developers/about/terms-of-service.html>
